Gradual Soundness: Lessons from Static Python

### 2022-01-18 pj-7.1/paper.tex
pj: draft running title + authors
rewrote soundness text

%\titlerunning{short title}

@@

%\authorrunning{K-C. Lu, B. Greenman, C. Meyer, D. Viehland, S. Krishnamurthi}

@@

To validate the design of \SP{}, we developed a model of the language in PLT
Redex~\cite{kcdeffmrtf-popl-2012}.
The model covers a substantial part of the Python language, including
module imports, loops, exception handlers, and delete statements.
It follows \SP{}'s approach to typing these features.
The model is available at the following URL:

\smallskip
{\centering\url{https://github.com/brownplt/insta-model}}

Because there is no prior formalization of Python and \SP{}, we cannot
\emph{verify} that the model matches reality.
Instead, we have applied the well-established method of testing end-to-end
conformance~\cite{gsk-ecoop-2010,gclpk-dls-2012,pmmwplck-oopsla-2013,bcfgmnss-popl-2014,fgpssmds-popl-2016}
with a test suite.
In particular, we translated \numSPtest{} tests from the \SP{} regression suite to the
syntax of the model and confirmed that the results
do match,\footnote{For most of the \numSPtest{} tests, the translation is automatic.
  A few tests required hand-pruning to remove features that the model
  does not handle (\totalnum{52}).}
which suggests that the model conforms to actual \SP{}.

Beyond the conformance suite, we have applied two further tests for correctness.
First, we used Redex's random testing tools~\cite{kf-sfp-2009} to check
type soundness on thousands of example programs (\totalnum{\mbox{FILL}}).
%% FILL how many programs exactly?
%% FILL what bugs did testing catch?
Second, we have manually inspected the model to convince ourselves
of its soundness.
Though not a proof, the inspection gives us confidence that the model,
and by extension \SP{}, is essentially sound.

>>>>

\titlerunning{DRAFT do not distribute}

@@

\authorrunning{DRAFT do not distribute}

@@

To validate the design of \SP{}, we developed a model of the language in PLT
Redex~\cite{kcdeffmrtf-popl-2012}.
The model covers a substantial part of the Python language, including
module imports, loops, exception handlers, and delete statements.
It follows \SP{}'s approach to typing these features.
The model is available at:
\begin{center}
  \url{https://github.com/brownplt/insta-model}
\end{center}
Because there is no prior formalization of (all of) Python and of \SP{}, we cannot
\emph{verify} that the model matches them. Instead, we have applied
three methods to give confidence that our model matches reality.
First, we manually reviewed the model in depth---using our
own expertise---to look for non-conformance.  Second, 
we used Redex's random testing tools~\cite{kf-sfp-2009} to check
type soundness on thousands of example programs (\totalnum{\mbox{FILL}}).
%% FILL how many programs exactly?
%% FILL what bugs did testing catch?
Finally, we employed the well-established method of testing end-to-end
conformance~\cite{gsk-ecoop-2010,gclpk-dls-2012,pmmwplck-oopsla-2013,bcfgmnss-popl-2014,fgpssmds-popl-2016}
with a test suite.
In particular, we translated \numSPtest{} tests from the \SP{} regression suite to the
syntax of the model and confirmed that the results
do match,\footnote{For most of the \numSPtest{} tests, the translation is automatic.
  A few tests required hand-pruning to remove features that the model
  does not handle (\totalnum{52}).}
which suggests that the model conforms to actual \SP{}.
